---
title: "The mnimum graph coloring problem"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The mnimum graph coloring problem}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

In this article we will model the [minimum graph coloring problem](https://en.wikipedia.org/wiki/Graph_coloring). The goal: color a map with as few colors as possible while no two adjacent regions having the same color.

## Read spatial data

First, let's load some useful packages needed for (spatial) data processing.

```{r, message=FALSE, warning=FALSE}
library(rgeos)
library(rgdal)
library(maptools)
library(dplyr)
```

Then we read in the 50 states of the US.

```{r}
# data from https://github.com/datasets/geo-boundaries-us-110m
# reference data from here: https://github.com/nvkelso/natural-earth-vector/tree/master/110m_cultural
# License of reference data: public domain (https://github.com/nvkelso/natural-earth-vector/blob/master/LICENSE.md)
map_data <- rgdal::readOGR("https://raw.githubusercontent.com/datasets/geo-boundaries-us-110m/84e946f6b1de01e2642bcdb17d5b697acb6b48c4/json/ne_110m_admin_1_states_provinces_shp_scale_rank.geojson")
```

Next step is to create an adjacency list to determine neighboring states.

```{r}
# this gives as an adjancy list
neighbors <- spdep::poly2nb(map_data)

# a helper function that determines if two nodes are adjacent
is_adjacent <- function(i, j) {
  purrr::map2_lgl(i, j, ~ .y %in% neighbors[[.x]])
}
```

We can then ask:
```{r}
is_adjacent(1, 2)
is_adjacent(1, 3)
```


## Optimization model

Next, we will model the problem with `rmpk` as a mixed integer linear program that tries to find a coloring with as few colors as possible.

```{r, message=FALSE}
library(rmpk)
library(ROI.plugin.glpk)
```

```{r, message=FALSE}
n <- nrow(map_data@data) # number of nodes aka states
max_colors <- 4 # 4 should be enough. But you increase this number.

# based on the formulation from here
# http://wwwhome.math.utwente.nl/~uetzm/do/IP-FKS.pdf
model <- MIPModel(ROI_solver("glpk", control = list(presolve = TRUE, verbose = TRUE)))
  
# 1 iff node i has color k
model$add_variable(x[i, k], type = "integer", i = 1:n, k = 1:max_colors, lb = 0, ub = 1)
  
# 1 iff color k is used
model$add_variable(y[k], type = "binary", k = 1:max_colors)
  
# minimize colors
# multiply by k for symmetrie breaking (signifcant diff. in solution time)
model$set_objective(sum_expr(k * y[k], k = 1:max_colors), sense = "max")
  
# each node is colored
model$add_constraint(sum_expr(x[i, k], k = 1:max_colors) == 1, i = 1:n)
  
# if a color k is used, set y[k] to 1
model$add_constraint(x[i, k] <= y[k], i = 1:n, k = 1:max_colors)
  
# no adjacent nodes have the same color
for (i in 1:n) {
  for (j in 1:n) {
    if (is_adjacent(i, j)) {
      model$add_constraint(x[i, k] + x[j, k] <= 1, k = 1:max_colors)
    }
  }
}

# we could also set some states to specific colors and see if
# that still produces a valid coloring.
model$set_bounds(x[1, 1], lb = 1)
model$set_bounds(x[2, 2], lb = 1)
model$set_bounds(x[3, 3], lb = 1)
model$set_bounds(x[4, 4], lb = 1)
```

```{r}
model
```


## Solve it

```{r, warning=FALSE}
model$optimize()
```

Yay. We found the minimal coloring with `r model$objective_value()` colors.

## Plot the result

Last step is to plot the result. First we will get the colors from the optimal solution.

```{r}
assigned_colors <- model$get_variable_value(x[i, k]) %>% 
  filter(value > 0.99) %>% 
  arrange(i)
```

```{r}
head(assigned_colors, 5)
```

Then we need to prepare the data for ggplot and join the colors to the data.

```{r, message=FALSE, warning=FALSE}
library(ggplot2)
color_data <- map_data@data
color_data$color <- assigned_colors$k
plot_data_fort <- fortify(map_data, region = "adm1_code") %>% 
  left_join(select(color_data, adm1_code, color), 
            by = c("id" = "adm1_code")) %>% 
  mutate(color = factor(color))
```

Now we have everything to plot it:

```{r}
ggplot(plot_data_fort, aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = color)) + 
  coord_quickmap() + 
  viridis::scale_fill_viridis(discrete = TRUE, option = "D")
```

## Feedback

Do you have any questions, ideas, comments? Or did you find a mistake? Let's discuss on [Github](https://github.com/dirkschumacher/rmpk/issues).
